Generate Real and Pseudo-absences

### 2023-12-07 scripts/download_dependent_variables.Rmd
generation of raw absences

aim <- st_read(file.path(base_p, 'AIM', 'AIM_records.shp'), quiet = TRUE) %>% 
  mutate(date = as.Date(DatVstd)) %>% 
  select(taxon, date) %>% 
  st_transform(4326)

@@

thinned <- lapply(splicies, ONEperCELL)
thinned <- bind_rows(thinned)
write.csv(thinned, '../data/raw/occurrence/thinned_occurrences.csv', row.names = F)
repeated_occurrences <- lapply(splicies, multiplesPERcell)
repeated_occurrences <- purrr::discard(repeated_occurrences, ~nrow(.) < 2)
repeated_occurrences <- bind_rows(repeated_occurrences)
write.csv(repeated_occurrences, '../data/raw/occurrence/repeated_occurrences.csv', row.names = F)

@@

>>>>

aim <- st_read(file.path(base_p, 'AIM', 'AIM_records.shp'), quiet = TRUE) %>% 
  mutate(date = as.Date(DatVstd)) %>% 
  select(taxon, date, PrmryKy) %>% 
  st_transform(4326)

@@

thinned <- lapply(splicies, ONEperCELL)
thinned <- bind_rows(thinned)
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences.shp', append = F)
repeated_occurrences <- lapply(splicies, multiplesPERcell)
repeated_occurrences <- purrr::discard(repeated_occurrences, ~nrow(.) < 2)
repeated_occurrences <- bind_rows(repeated_occurrences)
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences.shp', append = F)

@@

```{r remove points outside analytical area}
library(terra)
domain <- rast(nrows = 1, ncols = 1) # create a big empty raster, you can go in through sf too. 
ext(domain) <- c( -125.5, -100, 27,  50) # set the extent
crs(domain) <- "EPSG:4326" # define the projection

domain <- as.polygons(domain) |>  # convert to vector data
  st_as_sf() |> # convert to simple feature
  st_transform(5070)

thinned <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE)
thinned <- thinned[lengths(st_disjoint(thinned, domain)) ==  0, ]
st_write(thinned, '../data/raw/occurrence/thinned/thinned_occurrences.shp', append = FALSE)

repeated_occurrences <- st_read('../data/raw/occurrence/repeated/repeated_occurrences.shp', quiet = TRUE)
repeated_occurrences <- thinned[lengths(st_disjoint(repeated_occurrences, domain)) ==  0, ]
st_write(repeated_occurrences, '../data/raw/occurrence/repeated/repeated_occurrences.shp', append = FALSE)
```

### 2023-12-07 scripts/generate_absences.Rmd
generation of raw absences

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
blm_surf <- st_read(paste0(pout, 'coarse_surface.shp'))
blm_surf <- st_simplify(blm_surf, dTolerance = 500) # just make it a bit smaller. 
blm_surf <- st_make_valid(blm_surf)
```

```{r}

aim <- st_read('../data/raw/occurrence/AIM/AIM_records.shp') %>% 
  filter(SpcsLst == 'ACHY')

bbox <- absence_drawer(aim, bg_abs = 0.15)

```

>>>>

pout <- '/media/steppe/hdd/SDM_restorations/data/raw/coarse_blm/'
blm_surf <- st_read(paste0(pout, 'coarse_surface.shp'), quiet = TRUE)
blm_surf <- st_simplify(blm_surf, dTolerance = 500) # just make it a bit smaller. 
blm_surf <- st_make_valid(blm_surf)
```

```{r generate raw absences, eval = F}

records <- st_read('../data/raw/occurrence/thinned/thinned_occurrences.shp', quiet = TRUE)
records <- records %>% 
  group_by(taxon) %>% 
  filter(n() >= 15)
records <- split(records, f = records$taxon)

out <- lapply(records, absence_drawer, bg_abs = 0.15)
out <- bind_rows(out)
st_write(out, '../data/raw/occurrence/absences.shp')
```
